---
title: "Class notest 3.24"
format: html
embed-resources: true
---

```{r}
#| message: false

library(tidyverse)
library(modeldata)

movies <- matrix(c(38, 30, 24, 10, 8,
                   15, 17, 16, 9, 5,
                   12, 11, 19, 25, 13),
                 nrow = 3,
                 byrow = TRUE)
```

So far, we've considered a single variable, either quantitative (`t.test`) or categorical (`chisq.test`). Up next, two variables. Are they independent?

Example, `island` and `species` in the `penguins` set. Both of these variables are categorical, so it's natural to make a *contingency table*.

```{r}
table(penguins$species, 
      penguins$island)
```

Null: the variables are independent. Knowing `island` doesn't tell you anything about `species`.

Alternative: the variables are not independent. Knowing `island` changes the proportions of `species`.

```{r}
ggplot(penguins, 
       aes(x = island, 
           fill = species)) +
  geom_bar(position = "fill") +
  scale_fill_brewer(palette = "Dark2")
```

We test these hypotheses using `chisq.test` function, just passing it the contingency table.

```{r}
penguins_table <- table(penguins$species, 
                        penguins$island)
chisq.test(penguins_table)
```

The p-value is infinitesimal. This data provides strong evidence against the null. We conclude that `species` and `island` are not independent, each gives information about the other.

## Another example

```{r}
glimpse(stackoverflow)
```

I'm interested in the relationship between company size and country. While size is quantitative, there is a problem. 

```{r}
ggplot(stackoverflow, 
       aes(x = CompanySizeNumber)) +
  geom_histogram()
```

I'll group respondants as working either at a large company or a small one. I want a new column of T/F, called "large".

```{r}
stackoverflow <- stackoverflow %>% 
  mutate(large = (CompanySizeNumber >= 5000))

stack_table <- table(stackoverflow$Country, 
                     stackoverflow$large)
```

$H_0:$ `Country` and `large` are independent.

$H_a$: they are not independent.

(need a plot)

```{r}
chisq.test(stack_table)
```

The p-value is infinitesimal (again). We conclude country and large give information about one another. 


